---
title: "Exercise 1"
author: Sam Harper
format: html
---
```{r setup, message=F, echo=F}
pacman::p_load(here, tidyverse, did,
  haven, kableExtra, modelsummary)
```

## Question 1

We will start by computing group-time average treatment effects without including any covariates in the parallel trends assumption.

```{r data}
d <- read_dta(here("data", 
  "job_displacement_data.dta"))
glimpse(d)
kbl(head(d))
```

1)  Use the `did` package to compute all available group-time average treatment effects.

```{r q1_1, message = F, warning = F}
gt_atts <- did::att_gt(
  yname = "income",
  idname = "id",
  tname = "year",
  gname = "group",
  control_group = "nevertreated",
  base_period = "universal",
  data = d)

modelsummary(gt_atts)
```

2)  *Bonus Question* Try to manually calculate $ATT(g=1992, t=1992)$.
    Can you calculate exactly the same number as in part (a)?
    
The estimate in part (a) was -8990.85. This can be calculated by hand by restricting the sample to the group first treated in 1992 and the untreated group, as well as limiting to the years 1991 and 1992. We take the long difference in average income between 1991 and 1992 for the two groups (1992 and never treated) and then the difference between those long differences should be the estimate from part (a):

```{r q1_2, message = F}   
d %>% filter((group==1992 | group==0) & 
  (year==1991 | year==1992)) %>%
  mutate(g1992_1992 = if_else(group==1992, 
    "1:Treated", "0:Untreated")) %>%
  group_by(g1992_1992, year) %>%
  summarise(mincome = mean(income)) %>%
  select(g1992_1992, year, mincome) %>%
  pivot_wider(names_from = year, values_from = mincome,
    names_prefix = "year") %>%
  mutate(`Long diff` = `year1992` - `year1991`) %>%
  group_by() %>%
  mutate(ATT_1992_1992 = `Long diff` - lag(`Long diff`,
    default = NA)) %>%
  kable(digits = 2) %>%
  column_spec(5, bold = TRUE) %>%
  kable_classic(html_font = "Helvetica", full_width = F)
```

3)  Aggregate the group-time average treatment effects into an event
    study and plot the results. What do you notice? Is there evidence
    against parallel trends?
    
```{r q1_3}
es <- did::aggte(gt_atts, type = "dynamic")
ggdid(es)
```

Most of the pre-trends are consistent with a null impact; however, all of the point estimates are positive (indicating earnings may have been trending upward among the treated before the treatment, and the period 2 years before the treatment is inconsistent with the null. It may be possible to reject the null hypothesis since at least one of the pre-periods seems to be inconsistent with the null. However, overall I'm not sure I would argue that there is *strong* evidence to reject the parallel trends assumption.

4)  Aggregate the group-time average treatment effects into a single
    overall treatment effect. How do you interpret the results?

```{r q1_4}
did::aggte(gt_atts, type = "simple")
```
If we take the simple average of all of the group-time ATTs, this estimate suggest that job displacement reduces average income by \$4686, with a 95% confidence interval of \$-9562 to $189. 

# Question 2

A major issue in the job displacement literature concerns a version of
anticipation. In particular, there is some empirical evidence that
earnings of displaced workers start to decline *before* they are
actually displaced (a rough explanation is that firms where there are
mass layoffs typically “struggle” in the time period before the mass
layoff actually takes place and this can lead to slower income growth
for workers at those firms).

1)  Is there evidence of anticipation in your results from Question 1?

The event study plot in Question 1 makes it difficult to know whether or not there is evidence of anticipation, since the plot fixes the estimate at 1-year prior to the intervention to be zero. However, as noted above, there was some evidence in the pre-trends that could be suggestive of anticipation effects, though these mainly showed slightly larger earnings trends for those eventually treated.

2)  Repeat parts (a)-(d) of Question 1 allowing for one year of
    anticipation.

```{r q2_1, message = F, warning = F}
gt_atts_a <- did::att_gt(
  yname = "income",
  idname = "id",
  tname = "year",
  gname = "group",
  control_group = "nevertreated",
  base_period = "universal",
  anticipation = 1,
  data = d)

did::aggte(gt_atts_a, type = "simple")
```

The simple average ATT when allowing for 1-year of anticipation produces an estimate of \$-5284 with a 95% CI of -9826 to -742. Overall allowing for 1-year of anticipation produces an estimate that is further from the null than when not allowing for anticipation, though the two estimates are unlikely to be statistically (or economically) distinguishable.

```{r q2_3, message = F}
es_a <- did::aggte(gt_atts_a, type = "dynamic")
ggdid(es_a)
```

If we allow for a 1-year anticipatory effect, the resulting event study plot shows some limited evidence that is consistent with an anticipation story. The estimate for the year just prior to the intervention is a reduction of earnings of \$-3839. The confidence bounds on this estimate show that the data are consistent with a reduction of up to \$8792, but also with a potential increase of \$1105.  

# Question 3

Now, let’s suppose that we think that parallel trends holds only after
we condition on a person sex and race (in reality, you could think of
including many other variables in the parallel trends assumption, but
let’s just keep it simple). In my view, I think allowing for
anticipation is desirable in this setting too, so let’s keep allowing
for one year of anticipation.

1)  Answer parts (a), (c), and (d) of Question 1 but including `sex` and
    `white` as covariates.
    
```{r q3_1, message = F, warning = F}
gt_atts_c <- did::att_gt(
  yname = "income",
  idname = "id",
  tname = "year",
  gname = "group",
  xformla = ~female + white,
  control_group = "nevertreated",
  base_period = "universal",
  data = d)

did::aggte(gt_atts_c, type = "simple")
```



2)  By default, the `did` package uses the doubly robust approach that
    we discussed during our session. How do the results change if you
    use a regression approach or propensity score re-weighting?

<div style="display: none;">

</div>

# Question 4

Finally, the data that we have contains a variable called `occ_score`
which is roughly a variable that measures the occupation “quality”.
Suppose that we (i) are interested in including a person’s occupation in
the parallel trends assumption, (ii) are satisfied that `occ_score`
sufficiently summarizes a person’s occupation, but (iii) are worried
that a person’s occupation is a “bad control” (in the sense that it
could be affected by the treatment).

1)  Repeat parts (a), (c), and (d) of Question 1 but including
    `occ_score` in the parallel trends assumption. Continue to allow for
    1 year of anticipation effects.

<div style="display: none;">

</div>

2)  What additional assumptions (with respect to occupation) do you need
    to make in order to rationalize this approach?