---
title: "Exercise 1"
author: Sam Harper
format: html
---
```{r setup, message=F, echo=F}
pacman::p_load(here, tidyverse, did,
  haven, kableExtra, modelsummary)
```

## Question 1

We will start by computing group-time average treatment effects without including any covariates in the parallel trends assumption.

```{r data}
d <- read_dta(here("data", 
  "job_displacement_data.dta"))
glimpse(d)
kbl(head(d))
```

1)  Use the `did` package to compute all available group-time average treatment effects.

```{r gt_atts}
gt_atts <- did::att_gt(
  yname = "income",
  idname = "id",
  tname = "year",
  gname = "group",
  control_group = "nevertreated",
  base_period = "universal",
  data = d)

modelsummary(gt_atts)
```


# callaway and sant'anna including covariates
cs_x <- att_gt(yname="lemp",
               tname="year",
               idname="id",
               gname="G",
               xformla=~region,
               control_group="nevertreated",
               base_period="universal",
               data=data2)
cs_x_res <- aggte(cs_x, type="group")
summary(cs_x_res)
cs_x_dyn <- aggte(cs_x, type="dynamic")
ggdid(cs_x_dyn)
